Fixing Vulnerabilities Potentially Hinders Maintainability

### 2020-05-29 paper/icsme20.tex
fixes

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in 
previous work~\cite{Olivari:2018, cruz2019energyoriented}. Results show evidence of a 
trade-off between security and maintainability ($41.9\%$ of security patches yielded a negative impact), i.e., 
developers may be hindering software maintainability or even introduces 
new vulnerabilities while patching older ones. We also observe that some guidelines 
and programming languages are more likely to be affected than others. The implications 
of our study are that changes to codebases while patching 
vulnerabilities need to be performed with extra care (e.g., integrating tools
for patch risk assessment in the CI/CD pipelines); computer science
curricula needs to be updated; and, more secure programming languages are necessary.

>>>>

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in 
previous work~\cite{Olivari:2018, cruz2019energyoriented}. Results show evidence of a 
trade-off between security and maintainability, as $41.9\%$ of security patches yielded a negative impact. Hence,
developers may be hindering software maintainability while patching vulnerabilities. We also observe that some guidelines 
and programming languages are more likely to be affected than others. The implications 
of our study are that changes to codebases while patching 
vulnerabilities need to be performed with extra care (e.g., integrating tools
for patch risk assessment in the CI/CD pipelines); computer science
curricula needs to be updated; and, more secure programming languages are necessary.
